Baseline centering problem

### 2018-12-17 paper/tangotd.tex

Of course we cannot directly learn $\green$ from \eqref{eq:greenrep},
since $\Delta^{-1}$ in the right-hand-side is exactly what we want to
learn. But $\Delta^{-1}$ is the unique solution to the Bellman equations \eqref{eq:greenbell}; after
applying \eqref{eq:greenbell} to an arbitrary state $s_1$ and test
function $\psi$, and
substituting the model $\green$ for $\Delta^{-1}$ via
$\eqref{eq:greenrep}$, the left Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,P\Delta^{-1}\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s'\sim P(s'|s_1)}\green(s')\cdot
\E_{s_2\sim \rho}\,
f(s_2)\psi(s_2)
\end{gather}
because $\transp{\1_{s_1}}\,P=\E_{s'\sim P(s'|s_1)}\transp{\1_{s'}}\,$.
Meanwhile,
the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \green(s_1)\cdot\E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s'_2)$).

@@

Of course, we should apply this to relevant test functions $\psi$,
such as the reward function or any useful (predictive) features, namely,
any function we're interested in predicting the future values of. 
If just using a
random function $\psi$, there is very little signal to train
from. \footnote{An extreme case is to take a purely random $\psi$ at each
training step; then $\psi(s_1)$ and $\psi(s_2)$ are independent unless $s_1=s_2$,
which is a very unlikely event. This amounts to training with white noise
features. Formally it works thanks to the rare case $s_1=s_2$, but
variance is just terrible.}

Thus we may learn $\green$ via TD-like algorithms. We may even have
double sampling that works, thanks to featurization (using test functions
$\psi$), at least for the right equation. TODO should we featurize on the
left as well? Or train only with the right equation?

@@

The value function is obtained with $\psi=R$, namely,
\begin{equation}
V(s_1)=\transp{\1_{s_1}}\Delta^{-1}R\approx
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)R(s_2)
\end{equation}
Of course it is not convenient to have to sample from $s_2$. Instead one
might want to just learn the representation $f(R)$, defined by abuse of
notation as
\begin{equation}
f(R)\deq \E_{s_2\sim \rho} \,f(s_2)R(s_2)
\end{equation}
(which can be computed online)
and then
\begin{equation}
V(s_1)\approx \green(s_1)\cdot f(R)
\end{equation}

>>>>

Of course we cannot directly learn $\green$ from \eqref{eq:greenrep},
since $\Delta^{-1}$ in the right-hand-side is exactly what we want to
learn. But $\Delta^{-1}$ is the unique solution to the Bellman equations \eqref{eq:greenbell}; after
applying \eqref{eq:greenbell} to an arbitrary state $s_1$ and test
function $\psi$, and
substituting the model $\green$ for $\Delta^{-1}$ via
$\eqref{eq:greenrep}$, the left Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,P\Delta^{-1}\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s'\sim P(s'|s_1)}\green(s')\cdot
\E_{s_2\sim \rho}\,
f(s_2)\psi(s_2)
\end{gather}
because $\transp{\1_{s_1}}\,P=\E_{s'\sim P(s'|s_1)}\transp{\1_{s'}}\,$.
For a fixed $f$, 
this is basically a multiple Bellman equation on $\green$ with multiple
rewards $\psi$, and features $f\psi$.

Meanwhile,
the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \green(s_1)\cdot\E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s'_2)$).

@@

Of course, we should apply this to relevant test functions $\psi$,
such as the reward function or any useful (predictive) features, namely,
any function we're interested in predicting the future values of. 
If just using a
random function $\psi$, there is very little signal to train
from. \footnote{An extreme case is to take a purely random $\psi$ at each
training step; then $\psi(s_1)$ and $\psi(s_2)$ are independent unless $s_1=s_2$,
which is a very unlikely event. This amounts to training with white noise
features. Formally it works thanks to the rare case $s_1=s_2$, but
variance is just terrible.} Even training only on $\psi=R$ should be
interesting: the left Bellman equation just becomes a particular
parameterization of the value function as $V(s_1)=\green(s_1)\cdot
\E_{s_2} R(s_2)f(s_2)$, so if the left Bellman equation is used, this
will approximate the true $V$.

Thus we may learn $\green$ and $f$ via TD-like algorithms. (Does it make
sense to use the left equation for $\green$ and the right equation for
$f$?) We may even have
double sampling that works, thanks to featurization (using test functions
$\psi$), at least for the right equation. TODO should we featurize on the
left as well? Or train only with the right equation?

@@

The value function is obtained with $\psi=R$, namely,
\begin{equation}
V(s_1)=\transp{\1_{s_1}}\Delta^{-1}R\approx
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)R(s_2)
\end{equation}
Of course it is not convenient to have to sample from $s_2$. Instead one
might want to just learn the representation $f(R)$, defined by abuse of
notation as
\begin{equation}
f(R)\deq \E_{s_2\sim \rho} \,f(s_2)R(s_2)
\end{equation}
(which can be computed online)
and then
\begin{equation}
V(s_1)\approx \green(s_1)\cdot f(R)
\end{equation}
and of course it might be helpful to learn the remainder $V-\green\cdot
f(R)$ via ordinary TD, in case $R$ has no component along the features
learned.
